---
title: "House, with States"
author: Cameron Greenwalt, John Graff, Mitchell Rands
output: html_document
---

<style type="text/css">
h1.title {
font-size: 40px;
text-align: center;
}
h4.author {
font-size: 40px;
text-align: center;
}
</style>

```{r setup, include=FALSE}
library(tidyverse)
library(GGally)
library(ggfortify)  # plot lm objects using ggplot instead of base R
library(car)  # needed for added-variable plots and dfbetas and dffits
library(corrplot)  # colored correlation matrix
library(bestglm)
library(glmnet)
library(leaps)
library(fastDummies) # Easily makes dummy variables
library(stringr)
```

Load the dataset.
```{r}
house <- read.csv("house.csv", header = TRUE, sep = ",")
```


Since all houses in this dataset come from Washington, we will drop the statezip and country columns. This analysis will use only city as a predictor in house price, so the street column will be removed as well.
```{r}

house <- subset(house, select = c(price, bedrooms, bathrooms, sqft_living,
                                  sqft_lot, floors, waterfront, view,
                                  condition, sqft_above, sqft_basement,
                                  yr_built, yr_renovated, city))
```

Change the appropriate columns to factors.
```{r}
house$waterfront <- ifelse(house$waterfront == 0, "no", "yes")
house$waterfront <- as.factor(house$waterfront)

house$view <- as.factor(house$view)

house$condition <- as.factor(house$condition)

house$yr_renovated <- ifelse(house$yr_renovated == 0, "no", "yes")
house$yr_renovated <- as.factor(house$yr_renovated)
names(house)[13] <- "renovated"

house$city <- as.factor(house$city)
```

Now that our data frame has been set up appropriately, we can print summaries to begin data exploration
```{r}
head(house)
summary(house)
```
Create a linear model for the data set.
```{r}
house.lm <- lm(price ~ . , data = house)
```

Perform variable selection to narrow down the number of predictors. Since an exhaustive method is computationally unreasonable, sequential replacement and backward selection will be run.



Backward Selection
```{r}
regsubsets.backward.out <-
    regsubsets(price ~ .,
               data = house,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "backward")
regsubsets.backward.out
```

```{r}
regsubsets.seqrep.out <-
    regsubsets(price ~ .,
               data = house,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "seqrep")
regsubsets.seqrep.out
```

Plot the BIC and Cp for the Sequential Replacement Method.
```{r}
summary.seqrep.out <- summary(regsubsets.backward.out)

plot(summary.seqrep.out$bic,
     xlab="Number of Predicors",
     ylab="BIC")

plot(summary.seqrep.out$cp, xlab='P (# of predictors + 1)', ylab='Cp')
abline(a=0,b=1)
```

Plot the BIC and Cp for the Backward Selection method.
```{r, fig.align='center'}
summary.backward.out <- summary(regsubsets.backward.out)

plot(summary.backward.out$bic,
     xlab="Number of Predicors",
     ylab="BIC")

plot(summary.backward.out$cp, xlab='P (# of predictors + 1)', ylab='Cp')
abline(a=0,b=1)
```




BIC and Cp outputs from either selection method are very similar. To have a balance between low model complexity and low Cp value, we will average BIC and Cp to find a decent model.


Find out which model had the lowest Cp
```{r, fig.align='center'}
average <- (summary.seqrep.out$cp + summary.seqrep.out$bic) / 2
plot(average, xlab='P (# of predictors + 1)', ylab='Average between BIC and Cp')
which.min(average)
```


The model with 20 predictors is the model with the lowest BIC that still lies under the Cp line. We will use that model
```{r, fig.align='center'}
as.data.frame(summary.seqrep.out$outmat[20, ])
```

The predictors to include in our model are bedrooms, bathrooms, sqft_living, waterfront, view, sqft_above, yr_built, and city.
```{r, fig.align='center'}
house.new <- subset(house, select = c(price, bedrooms, bathrooms, sqft_living,
                                      waterfront, view, sqft_above, yr_built,
                                      city))
head(house.new, 5)

house.new.lm <- lm(price ~ ., data = house.new)
summary(house.new.lm)
```


#### Begin visual EDA


Print a scatterplot matrix for the continuous variables.
```{r, fig.align='center'}
pairs(house.new[, c(1, 2, 3, 4, 7, 8)], pch = 20)
```
Print residuals vs predictors plots and residuals vs fitted values plot
```{r, fig.align='center'}

# Residuals vs predictors
ggplot(data = house.new, 
       mapping = aes(x = bedrooms, y = house.new.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.new, 
       mapping = aes(x = bathrooms, y = house.new.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.new, 
       mapping = aes(x = sqft_living, y = house.new.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.new, 
       mapping = aes(x = sqft_above, y = house.new.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.new, 
       mapping = aes(x = yr_built, y = house.new.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)


# Residuals vs fitted values
autoplot(house.new.lm,
         which = 1,
         ncol = 1,
         nrow = 1) + theme_bw() +
  theme(aspect.ratio = 1) +
  scale_y_continuous()
```



The residuals vs fitted values plot doesn't have event variance. Try logging the response variable to see if that helps.
```{r, fig.align='center'}
# Remove houses with a price of zero. Logging won't work otherwise.
house.log <- house.new[!(house.new$price == 0),]

house.log$price <- log(house.log$price)
house.log.lm <- lm(price ~ . , data = house.log)
summary(house.log.lm)
```
Print a scatterplot matrix for the continuous variables.
```{r, fig.align='center'}
pairs(house.log[, c(1, 2, 3, 4, 7, 8)], pch = 20)
```

Print residuals vs predictors plots and residuals vs fitted values plot
```{r, fig.align='center'}

# Residuals vs predictors
ggplot(data = house.log, 
       mapping = aes(x = bedrooms, y = house.log.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.log, 
       mapping = aes(x = bathrooms, y = house.log.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.log, 
       mapping = aes(x = sqft_living, y = house.log.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.log, 
       mapping = aes(x = sqft_above, y = house.log.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)

ggplot(data = house.log, 
       mapping = aes(x = yr_built, y = house.log.lm$residuals)) +
  geom_point(alpha = 0.2) +
  theme_bw() +
  ylab("Residuals") +
  theme(aspect.ratio = 1)


# Residuals vs fitted values
autoplot(house.log.lm,
         which = 1,
         ncol = 1,
         nrow = 1) + theme_bw() +
  theme(aspect.ratio = 1) +
  scale_y_continuous()
```

Check for multi-collinearity
```{r, fig.align='center'}
house.only.continuous <- subset(house.log, select = c(price, bedrooms,
                                                      bathrooms, sqft_living,
                                                      sqft_above, yr_built))

round(cor(house.only.continuous), 2)
corrplot(cor(house.only.continuous), type = "upper")


house.log.vifs <- vif(house.log.lm)
house.log.vifs
max(house.log.vifs)
mean(house.log.vifs)
```


Sqft_living and sqft_above have a multi-collinearity problem. Since it makes more sense to keep sqft_living, we will remove sqft_above
```{r, fig.align='center'}
house.log <- house.log[, -7]
```

Remake the linear model and check for multicollinearity again.
```{r, fig.align='center'}
house.log.lm <- lm(price ~ . , data = house.log)
summary(house.log.lm)

house.only.continuous <- subset(house.log, select = c(price, bedrooms,
                                                      bathrooms, sqft_living,
                                                      yr_built))

round(cor(house.only.continuous), 2)
corrplot(cor(house.only.continuous), type = "upper")


house.log.vifs <- vif(house.log.lm)
house.log.vifs
max(house.log.vifs)
mean(house.log.vifs)
```

### Other linear regression assumptions to be put here . . .



Influential points -> DFFITS
```{r, fig.align='center'}
house.log.dffits <- data.frame("dffits" = dffits(house.log.lm))
house.log.dffits$obs <- 1:length(house.log$price)

ggplot(data = house.log.dffits) +
  geom_point(mapping = aes(x = obs, y = abs(dffits))) +
  geom_hline(mapping = aes(yintercept = 2 * 
                             sqrt(length(house.log.lm$coefficients) / 
                                    length(obs))),
             color = "red", linetype = "dashed") +
  theme_bw() +
  theme(aspect.ratio = 1)

dffits <- house.log.dffits[abs(house.log.dffits$dffits)
                              > 2 * sqrt(length(house.log.lm$coefficients)
                                         / length(house.log.dffits$obs)), ]
dffits$dffits <- abs(dffits$dffits) # This makes all dffits positive so the next
                                      # operation is easier

head(dffits[order(-dffits$dffits), ]) # This code orders the dffits dataframe
                                        # in descending order by dffits
```
Remove observation 4346 and remake the linear model.
```{r, fig.align='center'}
house.log.new <- house.log[-c(4346),] 

house.log.new.lm <- lm(house.log.new)
```

Recheck DFFITS
```{r, fig.align='center'}
house.log.new.dffits <- data.frame("dffits" = dffits(house.log.new.lm))
house.log.new.dffits$obs <- 1:length(house.log.new$price)

ggplot(data = house.log.new.dffits) +
  geom_point(mapping = aes(x = obs, y = abs(dffits))) +
  geom_hline(mapping = aes(yintercept = 2 * 
                             sqrt(length(house.log.new.lm$coefficients) / 
                                    length(obs))),
             color = "red", linetype = "dashed") +
  theme_bw() +
  theme(aspect.ratio = 1)

dffits <- house.log.new.dffits[abs(house.log.new.dffits$dffits)
                              > 2 * sqrt(length(house.log.new.lm$coefficients)
                                         / length(house.log.new.dffits$obs)), ]
dffits$dffits <- abs(dffits$dffits) # This makes all dffits positive so the next
                                      # operation is easier

head(dffits[order(-dffits$dffits), ]) # This code orders the dffits dataframe
                                        # in descending order by dffits
```






Now we will check for interactions between variables.
```{r, fig.align='center'}
house.inter <- lm(price ~ bedrooms * bathrooms +
                    bedrooms * sqft_living +
                    bedrooms * waterfront +
                    bedrooms * view +
                    bedrooms * yr_built +
                    bedrooms * city +
                    bathrooms * sqft_living +
                    bathrooms * waterfront +
                    bathrooms * view +
                    bathrooms * yr_built +
                    bathrooms * city +
                    sqft_living * waterfront +
                    sqft_living * view +
                    sqft_living * yr_built +
                    sqft_living * city +
                    waterfront * view +
                    waterfront * yr_built +
                    waterfront * city +
                    view * yr_built +
                    view * city  +
                    yr_built + city,
                  data = house.log.new)
summary(house.inter)

# sqft_living yr_built, sqft_living view, bathrooms sqft_living, bedrooms yr_built, bedrooms view, bedrooms waterfront, bedrooms sqft_living, 
```
Only keep the significant interactions
```{r, fig.align='center'}
# sqft_living yr_built, sqft_living view, bathrooms sqft_living, bedrooms yr_built, bedrooms view, bedrooms waterfront, bedrooms sqft_living, 

house.inter <- lm(price ~ . +
                    sqft_living * yr_built +
                    sqft_living * view +
                    sqft_living * bathrooms +
                    sqft_living * bedrooms +
                    bedrooms * yr_built +
                    bedrooms * view +
                    bedrooms * waterfront,
                  data = house.log.new)

summary(house.inter)

```
Now that our model is done, predict a house.
```{r, fig.align='center'}
prediction <- 
  predict(house.inter,
          newdata=data.frame(bedrooms=5,
                             bathrooms=3,
                             sqft_living=4500,
                             waterfront="yes",
                             view="3",
                             yr_built=2015,
                             city="Seattle"),
          interval="confidence",
          level=.95)

exp(prediction) # Convert back from the log scale
```

```{r, fig.align='center'}

```

```{r, fig.align='center'}

```

```{r, fig.align='center'}

```

```{r, fig.align='center'}

```


